Java Based Memory Latency Simulator

### 2020-11-21 finalPaper/paper.tex
Fixed some errors

Rotatational latency of a hard drive is how long does it take a platter to spin, as after all it is a mechanical
component bounded by a certain speed.
According to \cite{diskLatency}

\begin{tabular}{|c|c|}
  \hline
  RPMs & Rotational Latency(ms)\\
  \hline
  5400 & 11 ms\\
  \hline
  7200 & 8 ms\\
  \hline
  10000 & 6 ms\\
  \hline
  15000 & 4 ms\\
  \hline
\end {tabular}

Seek time; which is the time it takes for the Read-Write to move up and down the platter.
The Seek time itself is comprised of a series of phase times. The first being the acceleration phase,
the read write arm starts to move, then comes the coasting phase when the arm is moving at full speed. This
is followed by the deceleration phase where the read write arm slows down. Then finally comes the
settling phase as the read write head stops right at the data it needs.
The average seek time of hardrives of size 300 GB to 1TB ranged from an average time of 4ms to 9ms.
(source: \cite{three} )

@@

\begin{tabular}{|l|rr|}
  \hline
  HDD & Cheetah 15K.5 & Barracuda\\
  \hline
  Capacity & 300 GB & 1 TB\\
  \hline
  RPM & 15,000 & 7,200\\
  \hline
  Average Seek & 4 ms & 9ms\\
  \hline
  Platters & 4 & 4\\
  \hline
  T I/O Random & 6 ms & 13.2 ms\\
  \hline
  T I/O Sequential & 800 ms & 900 ms\\
  \hline
\end {tabular}

\subsection*{In our simulation}

@@

As of 2012 these were some of the Latencies, that includes different operation on a low level
and at the high level.(Source: \cite {Latencies})

\begin{tabular}{|l|c|}
  \hline
  L1 cache reference & 0.5 ns\\
  \hline
  Branch mispredict & 5 ns\\
  \hline
  L2 cache reference & 7 ns\\
  \hline
  Mutex lock/unclock & 25 ns\\
  \hline
  Main memory reference & 100 ns\\
  \hline
  Compress 1k bytes with zippy & 3000 ns / 3 us\\
  \hline
  Send 1 KB over Gbps network & 10000 ns / 10 us\\
  \hline
  Read 4 KB randomly from memory & 150000 ns / 150 us\\
  \hline
  Read 1 MB sequentially from memory & 250000 ns / 250 us\\
  \hline
  Round trip within same data center & 500000 ns / 500 us\\
  \hline
  Read 1 MB sequentially from SSD & 1000 us / 1 ms\\
  \hline
  Disk Seek & 10000 us / 10 ms\\
  \hline
  Read 1 MB sequentially from the disk & 20000 us / 20 ms\\ 
  \hline    
\end {tabular}

Not all of this information, especially concerning network is that closely relevant to our field of study, but
we can certianly see the latencies that included reading, writing and referencing from the CPU caches, memory
and the drives had a huge difference in between them. The low level operations that took place within the
CPU were the quickest of all, referencing Caches and branching. Secondly referencing the RAM (7 ns) was slower than
referencing the the caches (0.5 ns and 7 ns) but alot faster that going to a and reading from disk which was almost a
20 million times slower than refernecing the L1 Cache and upto a hundred thousand times slower than referencing
the RAM.

>>>>

Rotatational latency of a hard drive is how long does it take a platter to spin, as after all it is a mechanical
component bounded by a certain speed.
According to \cite{diskLatency} 
\\
\begin{tabular}{|c|c|}
  \hline
  RPMs & Rotational Latency(ms)\\
  \hline
  5400 & 11 ms\\
  \hline
  7200 & 8 ms\\
  \hline
  10000 & 6 ms\\
  \hline
  15000 & 4 ms\\
  \hline
\end {tabular}
\\
Seek time; which is the time it takes for the Read-Write to move up and down the platter.
The Seek time itself is comprised of a series of phase times. The first being the acceleration phase,
the read write arm starts to move, then comes the coasting phase when the arm is moving at full speed. This
is followed by the deceleration phase where the read write arm slows down. Then finally comes the
settling phase as the read write head stops right at the data it needs.
The average seek time of hardrives of size 300 GB to 1TB ranged from an average time of 4ms to 9ms.
(source: \cite{three} )

@@

\\
\begin{tabular}{|l|rr|}
  \hline
  HDD & Cheetah 15K.5 & Barracuda\\
  \hline
  Capacity & 300 GB & 1 TB\\
  \hline
  RPM & 15,000 & 7,200\\
  \hline
  Average Seek & 4 ms & 9ms\\
  \hline
  Platters & 4 & 4\\
  \hline
  T I/O Random & 6 ms & 13.2 ms\\
  \hline
  T I/O Sequential & 800 ms & 900 ms\\
  \hline
\end {tabular}
\\
\subsection*{In our simulation}

@@

As of 2012 these were some of the Latencies, that includes different operation on a low level
and at the high level.(Source: \cite {Latencies})
\\
\begin{tabular}{|l|c|}
  \hline
  L1 cache reference & 0.5 ns\\
  \hline
  Branch mispredict & 5 ns\\
  \hline
  L2 cache reference & 7 ns\\
  \hline
  Mutex lock/unclock & 25 ns\\
  \hline
  Main memory reference & 100 ns\\
  \hline
  Compress 1k bytes with zippy & 3000 ns / 3 us\\
  \hline
  Send 1 KB over Gbps network & 10000 ns / 10 us\\
  \hline
  Read 4 KB randomly from memory & 150000 ns / 150 us\\
  \hline
  Read 1 MB sequentially from memory & 250000 ns / 250 us\\
  \hline
  Round trip within same data center & 500000 ns / 500 us\\
  \hline
  Read 1 MB sequentially from SSD & 1000 us / 1 ms\\
  \hline
  Disk Seek & 10000 us / 10 ms\\
  \hline
  Read 1 MB sequentially from the disk & 20000 us / 20 ms\\ 
  \hline    
\end {tabular}
\\
Not all of this information, especially concerning network is that closely relevant to our field of study, but
we can certianly see the latencies that included reading, writing and referencing from the CPU caches, memory
and the drives had a huge difference in between them. The low level operations that took place within the
CPU were the quickest of all, referencing Caches and branching. Secondly referencing the RAM (7 ns) was slower than
referencing the the caches (0.5 ns and 7 ns) but alot faster that going to a and reading from disk which was almost a
20 million times slower than refernecing the L1 Cache and upto a hundred thousand times slower than referencing
the RAM.
